Inferring community characteristics in labelled networks

### 2021-05-27 nips/derivations.tex
More hyp testing

\subsection{Hypothesis test on feature weights}

@@

Using the approximation, we can construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability. Assuming without loss of generality that
%
\begin{equation}
	p(|W_{ij}| > c) = p(W_{ij})
\end{equation}

>>>>

\subsection{Hypothesis test on feature weights}
\label{appdx:hyp-test}

@@

Using the approximation, we wish to construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability, where $c>0$. More specifically, we wish to determine:
%
\begin{equation}
	\begin{aligned}
		H_0&: |W_{ij}| \leq c \\
		H_1&: |W_{ij}| > c
	\end{aligned}
\end{equation}
%
We can consider the probabilities $p(W_{ij} > c)$ and $p(W_{ij} < c)$ separately. Starting with $p(W_{ij} > c)$ we can write:
%
\begin{align*}
	p(W_{ij} > c) &= p\left(Z > \frac{c - \hat{\mu}_{ij}}{\hat{\sigma}_{ij}}\right) \\
	&= p\left(-Z < \frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)\\
	&= \Phi\left(\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)
\end{align*}
%
Where $Z \sim \Gaussian(0,1)$ was introduced as a temporary variable and $\Phi(\cdot)$ is the standard Gaussian cdf. Now we introduce the variable $k$ to control the degree of significance of the result. By enforcing $k$ to be less than the argument of $\Phi$ we show that: 
%
\begin{align*}
	\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}} &> k \\
	\hat{\mu}_{ij} - k \hat{\sigma}_{ij} &> c
\end{align*}
%
As $\Phi(\cdot)$ is monotonically increasing, this result gives us that:
%
\begin{equation}
	p(W_{ij} > c) > \Phi(k) \iff \hat{\mu}_{ij} - k \hat{\sigma}_{ij} > c
	\label{eqn:hyp-test-positive}
\end{equation}
%
A similar argument can be made to show that:
%
\begin{equation}
	p(W_{ij} < -c) > \Phi(k) \iff \hat{\mu}_{ij} + k \hat{\sigma}_{ij} < -c
	\label{eqn:hyp-test-negative}
\end{equation}
%
Clearly equations \ref{eqn:hyp-test-positive} and \ref{eqn:hyp-test-negative} cannot both hold as $c>0$. Nevertheless, a necessary and sufficient condition for one of them holding is:
%
\begin{equation}
	(\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c) = \emptyset
	\label{eqn:hyp-test-empty}
\end{equation}
%
We therefore choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty}. This tells us with confidence greater than $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$. Note that we choose to accept $H_0$ for the case where the combined probabilities of $p(W_{ij} < -c) + p(W_{ij} > c) > \Phi(k)$, if equation \ref{eqn:hyp-test-empty} is not satisfied. This is because we wish to apply this for dimensionality reduction purposes. We only want to keep weights that have a confident prediction outside the null-space not those that succeed by summing the two extremes of their distribution. Note that this case is very rare empirically.

### 2021-05-27 nips/inference.tex
More hyp testing

Once we have the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$. Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$, we can define:
%
\begin{equation}
	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
\end{equation}
%
A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is constructed as in equation \ref{eqn:kept-feature-set}.
%
\begin{align}
	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
	\label{eqn:fij} \\
	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) \neq \emptyset \right\}
	\label{eqn:kept-feature-set}
\end{align}
%
Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices. If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}). $\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
%
\begin{equation}
	H_0: |W_{ij}| \leq c \qquad
	H_1: |W_{ij}| > c
	\label{eqn:hyp-test-discard}
\end{equation}
%
The multiplier $k$ determines the degree of significance of the result. However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method. Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$. We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$. This is summarised in equation \ref{eqn:c-star}. This approach is often preferred as it fixes the number of reduced dimensions.
%
\begin{equation}
	c^* = \argmax_{c>0} (c : |\Dcal'| = D', k=k_0)
	\label{eqn:c-star}
\end{equation}
%
For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}.

>>>>

Once we have the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$. Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$, we can define:
%
\begin{equation}
	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
\end{equation}
%
A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is constructed as in equation \ref{eqn:kept-feature-set}.
%
\begin{align}
	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
	\label{eqn:fij} \\
	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}
	\label{eqn:kept-feature-set}
\end{align}
%
Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices. If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}). $\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
%
\begin{equation}
	H_0: |W_{ij}| \leq c \qquad
	H_1: |W_{ij}| > c
	\label{eqn:hyp-test-discard}
\end{equation}
%
The multiplier $k$ in equation \ref{eqn:fij} determines the degree of significance of the result. Appendix \ref{appdx:hyp-test} explores the derivation in more detail. However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method. Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$. We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$. This is summarised in equation \ref{eqn:c-star}. This approach is often preferred as it fixes the number of reduced dimensions.
%
\begin{equation}
	c^* = \argmax_{c>0} (c : |\Dcal'| = D', k=k_0)
	\label{eqn:c-star}
\end{equation}
%
For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}.
